\textbf{ Short and long-term forecasts will require models incorporating different processes and covariates }

### 2019-04-22 doc/manuscript.tex
writing intro

\title{\textbf{Integrating spatial and temporal patterns \\ in ecological forecasts }}

@@

We love ecological forecasting yada yada. Common approaches use spatial patterns, temporal patterns, and process-level understanding (i.e., mechanistic models). Since we will never know all the mechanisms, we will inevitably need to rely on patterns. But they have limitations.

@@

Our goal is to capitalize on the complementary strengths of the spatial and temporal approaches. The idea is to take predictions from a spatial and a temporal model and combine them with a weighted average. The model weights can change over time: we expect the temporal model to dominate in the short-term, and the spatial model to dominate in the long-term. How rapidly that transition occurs depends on the rate of those ``slow" processes such as evolution and species turnover. We might learn something interesting by studying the rate at which the model weights change.

>>>>

\title{\textbf{ Short and long-term forecasts will require models incorporating different processes and covariates }}

@@

Recently, increased interest in ecological forecasting
Primary goal is applied: inform decisions (Clark 2001, Dietze et al. 2017, Mouquet et al. 2017)
Important secondary goal is to advance basic research (Houlahan et al. 2017)
Emerging consensus that we need to do this

How to do it is not so clear
Lots of different approaches
Lots of argument about the virtues of process-based models and phenomenological, or correlative models (refs?)
We often see basic research justified with the argument that improved process-level understanding will improve predictions, in fact we have made this
argument ourselves (cite our own stuff?)
It is easy to find examples of how extrapolation of linear relationships gets us in trouble (examples? refs?)
But in many cases simple phenomenological models make surprisingly good predictions (examples? refs?) and they can avoid the parameter error often associated
with complex process-based models

Pragmatism is the way forward: use the models that make the best predictions
This requires making testable predictions
Thus (Dietze et al 2017) emphasis on near-term forecasting (emphasize that we are on board?)
Longer-term forecasts harder to test, hindcasts may be best bet (Williams refs? Blois?)

What current literature and the argument for near term forecasting has not explicitly stated is that different models, incorporating different kinds of processes will
work better for short vs long-term forecasts. We demonstrate this using simulation models. Our analyses show that: 
1. For short-term forecasts, phenomenological time series approaches are hard to beat
2. Long-term forecasts require understanding slow processes such as evolutionary and ecological selection as well as dispersal
3. Different kinds of data reflect the operation of different processes:
	longitudinal data capture autocorrelation and responses of current assemblage;
	data across spatial gradients capture long-term outcome of selection and dispersal;
	as the time-scale of the forecast increases, the importance of these processes shifts.
4. A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.

\section*{Modeling approach}

Simulate effects of environmental change on a system with known dynamics
Assume fully mechanistic, process-based model is impossible
Fit two phenomenological models, each representing processes operating at different time scales, and make forecasts with them

@@

We show that responses to environmental change reflect both kinds of processes but at different time scales. We demonstrate that  by taking predictions from a spatial and a temporal model and combine them with a weighted average. The model weights can change over time: we expect the temporal model to dominate in the short-term, and the spatial model to dominate in the long-term. How rapidly that transition occurs depends on the rate of those ``slow" processes such as evolution and species turnover.
